### 2019-01-07 draft.tex
edited methods

We built a population genetic model to examine the role of sexual conflict in breaking down reinforcement. We assume two previously allopatric populations have diverged at a local adaptation locus and male compatibility locus. Upon secondary contact, the populations exchange migrants and form maladaptive hybrids, creating selection for the evolution of a female barrier. We are interested in the evolution of reproductive isolation in a focal population. There are three loci relevant to our focal population, all on the same chromosome: the female barrier locus, the male compatibility locus, and the local adaptation locus.

Viability selection operates at the local adaptation locus \textbf{L}. The \emph{L} allele is adaptive in population one and the \emph{$\ell$} allele is adaptive in population two; both alleles are initially fixed in the population in which they are favoured. Maladaptive alleles are selected against with selection strength $s_n$ in population $n$.  Selection acts multiplicatively. For example, hybrid offspring heterozygous at the local adaptation locus($L \ell$) experience a selection strength of $(1-s_n)$ while maladaptive homozygotes experience a selection strength of $(1-s_n)^2$. This ecological divergence gives rise to postzygotic (ecological) isolation between population one and population two that can be reinforced by assortative mating.

Assortative mating between these (incipient) species is governed by a female locus, \textbf{F}, expressed in female gametes, and a male compatibility locus, \textbf{M}, expressed in male gametes.  Females with the barrier allele, \emph{F}, block fertilization by males without the male compatibility allele, \emph{M} (Fig. \ref{fig:model}B).  Females with the alternative allele, \emph{f}, can be fertilized by males carrying either \emph{M} or \emph{m}.  We assume that the barrier allele is initially rare in our focal population and absent in the non-focal population. Each generation begins with migration of haploid males.  Formation of diploid zygotes is governed by the assortative mating loci.  Diploid offspring undergo viability selection before forming haploid gametes via meiosis (involving both recombination and segregation). We track the change in genotype frequencies after each generation to understand the evolution of reproductive isolation.

While we describe $F$ as the `barrier' allele throughout, it is important to note that $F$ is only an effective barrier to hybridization if the compatibility allele ($M$) is not common in the other species.  Therefore, reproductive isolation is measured as the fraction of heterospecific males that are rejected (\emph{Methods}, equation \ref{Eq:eq4}. Reinforcement is considered complete (equals one) when all heterospecific males are rejected.  This definition is useful for understanding the short- and long-term consequences of sexual conflict on the evolution of reproductive isolation.

Sexual conflict over hybridization arises between females and the fraction of males in the mating pool that are migrants.  Males that have migrated will be selected to hybridize in order to avoid associated opportunity costs, while females will be selected to avoid making unfit hybrid offspring, preferring conspecific males.  To measure the extent of the conflict we calculated female and migrated male fitness relative to their respective optimum fitness.  Female fitness is calculated as the absolute fitness (after mating and selection) of all diploid offspring in a population.  Optimum female fitness is the absolute fitness of diploid offspring if all heterospecific male gametes are rejected.  It is important to emphasize that optimum fitness will not always be one, but varies from generation to generation with the frequency of maladaptive alleles in the population.  Because the conflict is between resident females and \emph{immigrant} males, migrant male fitness is defined as the absolute fitness of only the diploid offspring sired by heterospecific males.  Optimum migrant male fitness is the absolute fitness if all immigrant males fertilized females (in terms of the conflict, males would have no fitness if all immigrant males were rejected).  Similar to optimum female fitness, optimum migrant male fitness varies with local adaptation allele frequency. Rigorously defining the conflict in this manner helps us understand the evolutionary forces driving or hindering reproductive isolation. It also allows us to design a model without sexual conflict.

We compare our model to one without sexual conflict by using a premating barrier instead of a postmating prezygotic barrier.  The key aspect of the premating model is that migrated males rejected by heterospecific females have the opportunity to `return' to their native population and fertilize conspecific females.  \textcolor{red}{EMPIRICAL EXAMPLES pollinator behaviour, flowering time, reproductive cycle (sea urchins), animals -> rejected by females and search until the find a mate that accepts them \textcolor{red}{Levin (1978), Grant (1994) from Moyle et al (2004)}}. Male and female interests are aligned because males that are rejected by heterospecific females (and therefore do not hybridize) have higher fitness than males that do hybridize by overcoming the premating isolating barrier.  We model this premating barrier by assuming that the migration rate is genotype dependent.  For example, $M f \ell$ male gametes will always have $\eta[i_1 i_2] = \eta_{max}$ whereas $M f \ell$ gametes will only have $\eta[i_1 i_2] = \eta_{max}$ if there is no barrier in the focal population, and $\eta[i_1 i_2] < \eta_{max}$ otherwise.

@@

We track haploid genotype frequencies using forward iterations of the recursion equations described in the \emph{Methods}.  With a total of six loci and 64 haploid genotypes in each species, the equations are sufficiently complicated that we could not obtain analytical solutions.  Furthermore, because the evolution of reinforcement is transient (see Fig. \ref{fig:sample}C), we are interested in the short-term dynamics of the system instead of long-term equilibria.

>>>>

We built a population genetic model to examine the role of sexual conflict in breaking down reinforcement. We assume two previously allopatric populations have diverged at a local adaptation locus and male compatibility locus. Upon secondary contact, the populations exchange migrants and form maladaptive hybrids, creating selection for the evolution of a female barrier. We are interested in the evolution of reproductive isolation in a focal population. There are three loci relevant to our focal population, all on the same chromosome: the female barrier locus, the male compatibility locus, and the local adaptation locus. We note that while our model of conflict over reinforcement is quite general, it was motivated by the biology of pollen-pistil incompatibility mediated by ``'gametophytic factors'' in hybridizing populations of \textit{Zea mays spp.}. In  \textit{Zea mays mexicana} male compatibility alleles are generally fixed and female barriers are common when sympatric with maize, but rare otherwise. By contrast, pollen compatibility and stylar barrier alleles are absent in allopatric populations of \textit{Zea mays mays} landraces although frequencies can vary dramatically in sympatry  \citep[see e.g. ][]{kermicle2006}.

Viability selection operates at the local adaptation locus \textbf{L}. The \emph{L} allele is locally adapted to population one's environment and the \emph{$\ell$} allele is locally adapted to population two's environment; both alleles are initially fixed in the population in which they are favoured. Maladapted alleles are selected against with selection strength $s_n$ in population $n$.  We assume that selection acts multiplicatively. For example, hybrid offspring heterozygous at the local adaptation locus($L \ell$) experience a selection strength of $(1-s_n)$ while maladaptive homozygotes experience a selection strength of $(1-s_n)^2$. This ecological divergence gives rise to postzygotic (ecological) isolation between population one and population two that can be reinforced by assortative mating.

Assortative mating between these (incipient) species is governed by a female locus, \textbf{F}, expressed in female gametes, and a male compatibility locus, \textbf{M}, expressed in male gametes.  Females with the barrier allele, \emph{F}, block fertilization by males without the male compatibility allele, \emph{M} (Fig. \ref{fig:model}B).  Females with the alternative allele, \emph{f}, can be fertilized by males carrying either \emph{M} or \emph{m}.  We assume that the barrier allele is initially rare in our focal population and absent in the non-focal population. Each generation begins with migration of haploid male gametes.  Formation of diploid zygotes is governed by these post-mating pre-zygotic isolation loci.  Diploid offspring undergo viability selection before forming haploid gametes via meiosis (involving both recombination and segregation). We track the change in genotype frequencies after each generation to understand the evolution of reproductive isolation.

We describe $F$ as the `barrier' allele throughout, but note that $F$ is only an effective barrier to hybridization if the compatibility allele ($M$) is rare or absent in the other species.  Therefore, reproductive isolation is measured as the fraction of migrant males that are rejected (\emph{Methods}, equation \ref{Eq:eq4}). 
% you said heterospecific.. I don't know what that means but I assumed it was "migrant"
Reinforcement is considered complete (equals one) when all migrant male gametes are rejected.  This definition is useful for understanding the short- and long-term consequences of sexual conflict on the evolution of reproductive isolation.

Sexual conflict over hybridization arises between females and the fraction of males in the mating pool that are migrants.  Under the conflict model, gametes have only one  reproductive opportunity, so migrant male gametes that can overcome interspecific barriers will be favored as they can only increase their fitness by hybridizing. By contrast, females that avoid the production of unfit hybrids will have more successful offspring than those who do not. To measure the extent of the conflict, we calculated female and migrated male fitness relative to their respective maximum potential fitness.  
% replacing optimum with maximum
Female fitness is calculated as the absolute fitness (after mating and selection) of all diploid offspring in a population.  Maximum potential  female fitness is the absolute fitness of diploid offspring if all migrant male gametes are rejected.  It is important to emphasize that optimum fitness will not always be one, but varies from generation to generation with the frequency of maladaptive alleles in the population.  Because the conflict is between resident females and \emph{immigrant} males, migrant male fitness is defined as the absolute fitness of only the diploid offspring sired by migrant male gametes.  Maximum potential migrant male gamete fitness is the absolute fitness if all immigrant males fertilized females (in terms of the conflict, males would have no fitness if all immigrant males were rejected).  Similar to maximum potential female fitness, maximum potential migrant male gamete fitness varies with local adaptation allele frequency. Rigorously defining the conflict in this manner helps us understand the evolutionary forces driving or hindering reproductive isolation. It also allows us to design a model without sexual conflict.

We compare our model to one without sexual conflict by using a premating barrier instead of a postmating prezygotic barrier.  The key aspect of the premating model is that migrated males rejected by migrant females have the opportunity to `return' to their native population and fertilize conspecific females. 
In addition to serving as an intellectually fair  control for our conflict model, this non-conflict model is potentially consistent with numerous empirical scenarios,  (\textcolor{red}{FOR EXAMPLE pollinator behaviour, flowering time, reproductive cycle (sea urchins), animals -> rejected by females and search until the find a mate that accepts them \textcolor{red}{Levin (1978), Grant (1994) from Moyle et al (2004). ALI CAN YOU CLEAN THIS UP.. I'M NOT IN YOUR HEAD ENOUGH}}. Male and female interests are aligned because migrant males that are rejected by heterospecific females (and therefore do not hybridize) have a second opportunity to mate with a conspecific and therefore have higher fitness than males that do hybridize by overcoming the premating isolating barrier.  To implement this non-conflict model, we model this premating barrier by assuming that the migration rate is genotype dependent.  For example, $M f \ell$ male gametes will always have $\eta[i_1 i_2] = \eta_{max}$ whereas $M f \ell$ gametes will only have $\eta[i_1 i_2] = \eta_{max}$ if there is no barrier in the focal population, and $\eta[i_1 i_2] < \eta_{max}$ otherwise.

@@

We track haploid genotype frequencies using forward iterations of the recursion equations described in the \emph{Methods}.  With a total of six loci and 64 haploid genotypes in each population, the equations are sufficiently complicated that we could not obtain analytical solutions.  Furthermore, because the evolution of reinforcement is transient (see Fig. \ref{fig:sample}C), we are interested in the short-term dynamics of the system instead of long-term equilibria.
